---
title: "R Notebook"
output: html_notebook
---

Clear all previously loaded packages
```{r}
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```

```{r}
library(data.table)
library(dplyr)
library(org.Hs.eg.db)
library(factoextra)
library(ggplot2)
library(stats)
library(ggpmisc)
library(ggforce)
library(tidyverse)
library(pheatmap)
library(viridis)
```

Load data
```{r}
term2gene_secM <- data.table(readRDS("Data/secMfeizi_term2gene.RDS"))
list.TPM_pheno_map <- readRDS("Output/list.TPM_pheno_map.RDS")
CHO96 <- list.TPM_pheno_map$pheno.data
```

############################################################
FUNCTIONS
############################################################

Correlation
```{r}
cor.test.hm <- function(x,y) {
  res <- cor.test(x,y, method="spearman")
  res.df <- data.frame(rho=res$estimate, pval=res$p.value)
  return(res.df)
}

geneCorr <- function(group.info, exp.data) {
  TPM.subset <- subset(exp.data, select=CHO96$sample_ID[CHO96$group.info %in% group.info])
  CHO96.subset <- CHO96[CHO96$sample_ID %in% colnames(TPM.subset) ,]
  if (!identical(colnames(TPM.subset), CHO96.subset$sample_ID)) {
    break
  }
  res.cor <- apply(TPM.subset, 1, cor.test.hm, y=CHO96.subset$Total.purified.protein_stock1.stock2_microg) %>% rbindlist(idcol=T, fill=T) %>%
  dplyr::rename("gene"=".id") %>% mutate(padj=p.adjust(pval, method="BH"))
}
```

Gene score
```{r}
geneScore <- function(TPM.matrix){
  threshold <- apply(TPM.matrix, 1, mean) %>% setNames(rownames(TPM.matrix))
  geneScore.matrix=matrix(, nrow=nrow(TPM.matrix), ncol=ncol(TPM.matrix))
  for (i in 1:nrow(TPM.matrix)){
    geneScore.matrix[i,] <- 5*log((1+TPM.matrix[i,])/threshold[i])
  }
  geneScore.matrix <- geneScore.matrix %>% magrittr::set_rownames(rownames(TPM.matrix)) %>% magrittr::set_colnames(colnames(TPM.matrix))
  return(geneScore.matrix)
}
```

correlation between activity scores and protein abundance
```{r}
subsystem.correlation <- function(group.info, scoreMatrix) {
  subsystemScore.subset <- scoreMatrix[, CHO96$group.info %in% group.info]
  CHO96.subset <- CHO96[CHO96$genes %in% colnames(subsystemScore.subset) ,]
  if (!identical(colnames(subsystemScore.subset), CHO96.subset$genes)) {
    break
  }
  res.cor <- apply(subsystemScore.subset, 1, cor.test.hm, y=CHO96.subset$Total.purified.protein_stock1.stock2_microg) %>% 
    rbindlist(idcol=T, fill=T) %>%
    dplyr::rename("gene"=".id") %>% mutate(padj=p.adjust(pval, method="BH"))
}
```

correlation within cluster
```{r}
geneCorr.cluster <- function(clust.id) {
  samples <- clust.samples$sample_ID[clust.samples$cluster==clust.id]
  TPM.subset <- TPM.log.secM[, samples]
  CHO96.subset <- CHO96[CHO96$sample_ID %in% colnames(TPM.subset) ,]
  if (!identical(colnames(TPM.subset), CHO96.subset$sample_ID)) {
    break
  }
  res.cor <- apply(TPM.subset, 1, cor.test.hm, y=CHO96.subset$Total.purified.protein_stock1.stock2_microg) %>%
    rbindlist(idcol=T, fill=T) %>%
    dplyr::rename("gene"=".id") %>% mutate(padj=p.adjust(pval, method="BH"))
}
```

plot significant correlations within each cluster
```{r}
sig.clust.cor <- function(clusterID, FDR.thresh, rho.thresh) {
  
  sig <- cluster.gene.corr[[clusterID]][padj <= FDR.thresh & abs(rho) >=rho.thresh,][order(-rho),] %>% 
  left_join(secM_TPM.map[, c("choSymbol", "Subsystem")], by=c("gene"="choSymbol")) %>% dplyr::rename(FDR=padj) %>%
  distinct()
  
  sig<- sig[order(Subsystem, -rho) ,][order(nrow(sig):1) ,]
  sig$gene <- factor(sig$gene, levels=sig$gene)
  
  p.cor <- ggplot(sig, aes(x=gene, y=rho)) +
  geom_segment( aes(x=gene, xend=gene, y=0, yend=rho, color=Subsystem), size=1.3, alpha=0.9) +
  geom_point(aes(color=Subsystem, size=FDR)) +
  coord_flip() +
  theme_bw() + #theme_light()
  ylab("Correlation with protein yield") +
  xlab("secM gene") +
  theme(text=element_text(size=17)) +
  scale_y_continuous(breaks = seq(-1, 1, by = .2))
  
  return(list(sig.df=sig, sig.plot=p.cor))
}
```

Anova by cluster
```{r}
anova.fxn <- function(x) {
  return(tryCatch(car::Anova(lm(paste(x, "~ cluster"), data=anova.feat)) %>% as.data.frame(), 
                  error=function(e) NULL))}
```

welch low vs high per cluster
```{r}
welch.cluster <- function(cluster.ID, exp.data, alt) {
  high.prod <- data.clust[cluster2==cluster.ID & amount_microgram=="highest amount", sample_ID]
  high.exp <- exp.data[, colnames(exp.data) %in% high.prod]
  low.prod <- data.clust[cluster2==cluster.ID & amount_microgram=="lowest amount", sample_ID]
  low.exp <- exp.data[, colnames(exp.data) %in% low.prod]
  
  keep <- matrixStats::rowSds(cbind(high.exp, low.exp)) >0.00000000001
  high.exp <- high.exp[keep ,]
  low.exp <- low.exp[keep ,]
  identical(rownames(high.exp), rownames(low.exp))
  
  res.df <- data.table(gene=rownames(high.exp), pval=rep(NA, nrow(high.exp)))
  for (i in 1:nrow(high.exp)) {
    welch.test <- t.test(low.exp[i,], high.exp[i,], alternative=alt)
    res.df$pval[i] <- welch.test$p.value
    res.df$mean.x[i] <- welch.test$estimate[1]
    res.df$mean.y[i] <- welch.test$estimate[2]
  }
  res.df <- res.df %>% mutate(padj=p.adjust(pval, method="BH"))
  
  return(res.df)
}
```

Custom color bar
```{r}
color.bar <- function(lut, min, max=-min, nticks=11, ticks=seq(min, max, len=nticks), title='') {
    scale = (length(lut)-1)/(max-min)

    dev.new(width=1.75, height=5)
    plot(c(0,10), c(min,max), type='n', bty='n', xaxt='n', xlab='', yaxt='n', ylab='', main=title,cex.main=1.8)
    axis(2, ticks, las=1, cex.axis=1.5)
    for (i in 1:(length(lut)-1)) {
     y = (i-1)/scale + min
     rect(0,y,10,y+1/scale, col=lut[i], border=NA)
    }
}
```

############################################################
secM genes missing in CHO (or at least not mappable)
############################################################

CHO.human map has already been trimmed to only include genes in TPM matrix
Dataframe of secM genes missing or not expressed in CHO:
```{r}
ori.map <- fread("ConversionTable/180405CHO_HUMAN_SYMBOL_ID_MAPPING_all.tsv") %>% .[, humanID:=as.character(humanID)]
ori.TPM <- readRDS("Data/txi.cho.RDS") %>% .[["abundance"]] 

# How many genes in our CHO tpm data is NOT mappable (not included in map): 6298/23706 (27%)
length(setdiff(rownames(ori.TPM), ori.map$choSymbol))

# How many human secM genes (feizi) are not in map: 16/581 (3%)
length(setdiff(term2gene_secM$ENTREZID, ori.map$humanID))
#View(term2gene_secM[term2gene_secM$ENTREZID %in% setdiff(term2gene_secM$ENTREZID, ori.map$humanID) ,])
length(setdiff(term2gene_secM$gene_name, ori.map$humanSymbol))

# subset map to only include secM genes
m.entrez <- ori.map[ori.map$humanID %in% term2gene_secM$ENTREZID ,] %>% left_join(term2gene_secM, by=c("humanID"="ENTREZID")) %>%
  mutate(ENTREZID=humanID)
m.symbol <- ori.map[ori.map$humanSymbol %in% term2gene_secM$gene_name ,] %>% left_join(term2gene_secM, by=c("humanSymbol"="gene_name")) %>%
  mutate(gene_name=humanSymbol)
secM.human_CHO_map <- rbind(m.entrez, m.symbol) %>% distinct()

# Subset map even further to only include genes in TPM data
length(setdiff(secM.human_CHO_map$choSymbol, rownames(ori.TPM))) #12, that map to 33 unique human genes
length(intersect(secM.human_CHO_map$choSymbol, rownames(ori.TPM))) #570 unique secM genes
secM_TPM.map <- secM.human_CHO_map[secM.human_CHO_map$choSymbol %in% rownames(ori.TPM) ,] #filter human-CHO map to only include genes in 

### discrepency between gene name in TPM, choSymbol, and secMs (aka have some TPM genes that match up with secM genes, however they are missing or hae a different alias in the map)
manual.map <- data.frame(choSymbol="Ufd1l", gene_name="UFD1L") %>% rbind(data.frame(choSymbol="Vimp", gene_name="VIMP")) %>% 
  rbind(data.frame(choSymbol="Park2", gene_name="PARK2")) %>% rbind(data.frame(choSymbol="Large", gene_name="LARGE")) %>%
  rbind(data.frame(choSymbol="Gyltl1b", gene_name="GYLTL1B")) %>% rbind(data.frame(choSymbol="Rfwd2", gene_name="RFWD2")) %>%
  left_join(term2gene_secM) %>% dplyr::rename(humanID=ENTREZID)
#add genes back into map
secM_TPM.map <- plyr::rbind.fill(secM_TPM.map, manual.map)

#####
missing.secM <- term2gene_secM[!(term2gene_secM$gene_name %in% secM_TPM.map$gene_name),]
#setdiff(missing.secM$ENTREZID, ori.map$humanID) %>% length()
setdiff(missing.secM$gene_name, ori.map$humanSymbol) # missing due to not being in map
intersect(missing.secM$gene_name, ori.map$humanSymbol) # present in map, but missing in TPM data


```
SUMMARY: There are 581 unique Human secM genes (581 unique ENTREZIDs map 575 unique gene symbols)
16 human secM genes are not found in our CHO-Human map; 9 secM genes from map can't be found in our TPM data
The result is a map of 556 unique secM genes (16+9+556=581) that map to 570 unique CHO genes
We lost a total of 25 secM genes (16 due to missing in map, 9 due to missing in TPM data)

Subset transcriptomics data to only include secMs (REMEMBER TO REMOVE BAD SAMPLES: "P6975_194", "P6975_108")
```{r}
TPM.secM <- ori.TPM[rownames(ori.TPM) %in% secM_TPM.map$choSymbol ,]
TPM.log.secM <- log2(TPM.secM + 0.001)
```

How many of the secM genes are not expressed or super lowly expressed?
```{r}
secM.0 <- secM_TPM.map[secM_TPM.map$choSymbol %in% rownames(TPM.secM)[matrixStats::rowSums2(TPM.secM) == 0] ,]
low.count <- apply(TPM.secM, 1, function(x) {sum(x<1)})
```

############################################################
Correaltion between secM expression and titers
############################################################
exclude WT and non producers
```{r}
# "highest cell count", "highest viability", "highest amount", "lowest amount", "lowest viability", "lowest cell count", "no amount", "WT" 
cor.genes <- geneCorr(group.info = c("highest cell count", "highest viability", "highest amount", "lowest amount", "lowest viability", "lowest cell count"), exp.data = TPM.log.secM) %>% left_join(secM_TPM.map[, c("choSymbol", "Subsystem")], by=c("gene"="choSymbol"))

summary(cor.genes$padj)
```
Result: no specific secM genes correlate with titers

############################################################
Calculate sec pathway activty scores (13 subsystems)
############################################################

```{r}
# remove zero genes
TPM.matrix <- TPM.secM[!(rownames(TPM.secM) %in% secM.0$choSymbol) ,]
```

```{r}
geneScore.matrix <- geneScore(TPM.matrix = TPM.matrix)
```

List of sample scores by subsystem
```{r}
#list of scores for each subsystem
subsystemScore.list=list()
for (subsystem in unique(term2gene_secM$Subsystem)) {
  subsystem.genes <- secM_TPM.map[secM_TPM.map$Subsystem==subsystem ,]
  geneScore.subset <- geneScore.matrix[rownames(geneScore.matrix) %in% subsystem.genes$choSymbol ,]
  subsystemScore.list[[subsystem]] <- colSums(geneScore.subset)/nrow(geneScore.subset)
}

#dataframe (rows=subsystems, cols=sample)
subsystemScore.df <- do.call("rbind", subsystemScore.list)
data.frame(avg=apply(subsystemScore.df, 1, mean), stdev=apply(subsystemScore.df, 1, stats::sd))
```

Visualization with pheatmap
```{r}
identical(colnames(subsystemScore.df), CHO96$sample_ID)
openxlsx::write.xlsx(as.data.frame(subsystemScore.df), "Output/secPathway_subsystem_scores.xlsx", rowNames=TRUE)
colnames(subsystemScore.df) <- CHO96$genes
row.ann <- data.frame(group.info=CHO96$group.info) %>% `rownames<-` (CHO96$genes)

##### Normalize change in pathway activity
WT <- subsystemScore.df[,colnames(subsystemScore.df)=="WT"] %>% setNames(rownames(subsystemScore.df)) %>%
  as.matrix
pathway.sds <- matrixStats::rowSds(as.matrix(subsystemScore.df[, 1:95])) %>% setNames(rownames(subsystemScore.df))

delta_subsystemScore <- apply(subsystemScore.df, 2, function(x) {x-WT}) %>% `rownames<-` (rownames(subsystemScore.df)) %>% 
  apply(2, function(x){x/pathway.sds})


###### plot
#row.ann.95 <- data.frame(group.info=CHO96$group.info[1:95]) %>% `rownames<-` (CHO96$genes[1:95])
row.ann.95 <- CHO96[1:95, c("amount_microgram", "conc", "viability")] %>% `rownames<-` (CHO96$genes[1:95]) %>% `colnames<-` (c("amount", "concentration", "viability"))
ann.col <- list(amount=c("highest amount"="#F8766D", "lowest amount"="#00BA38", "no amount"="#619CFF"), concentration=c("highest concentration"="slateblue4", "lowest concentration"="slateblue1"), viability=c("highest viability"="#d37703", "lowest viability"="#fcbb68"))
ann.col <- list(amount=c("highest amount"="#00356b", "lowest amount"="#66b2ff", "no amount"="red"), concentration=c("highest concentration"="#64064b", "lowest concentration"="#f890dc"), viability=c("highest viability"="#d37703", "lowest viability"="#fcbb68"))

row.ann.95 <- data.frame(amount=CHO96$amount_microgram[1:95]) %>% `rownames<-` (CHO96$genes[1:95]) 
ann.col <- list(amount=c("highest amount"="#F8766D", "lowest amount"="#00BA38", "no amount"="#619CFF"))

paletteLength=90
myColor <- colorRampPalette(c("navy", "white", "firebrick3"))(paletteLength)
myBreaks <- c(seq(min(delta_subsystemScore[,-96]), 0, length.out=ceiling(paletteLength/2) + 1), 
              seq(max(delta_subsystemScore[,-96])/paletteLength, max(delta_subsystemScore[,-96]), length.out=floor(paletteLength/2)))

## cluster by rows
png("Figures/Figure4a.png", width=550, height=800)
clust <- pheatmap(t(delta_subsystemScore[,-96]), fontsize=18, fontsize_col=15,
                               cluster_rows = T, cluster_cols = T, annotation_row = row.ann.95, gaps_col=T, cutree_rows=4, annotation_colors = ann.col, show_rownames=F, color=myColor, breaks=myBreaks, annotation_legend = F)
dev.off()

saveRDS(clust, "Figures/R_plots/Figure4a.RDS")

clust.samples <- cutree(clust$tree_row, k=4) %>% as.data.frame() %>% setNames("cluster") %>% rownames_to_column(var="gene")
identical(clust.samples$gene, CHO96$genes[1:95])
clust.samples <- clust.samples %>% mutate(sample_ID=CHO96$sample_ID[1:95], yield=CHO96$Total.purified.protein_stock1.stock2_microg[1:95])
```

Plot WT
```{r}
#png("Figures/Figure4a_WT.png", width=330, height=400)
pheatmap(cbind(WT[clust$tree_col$order,], WT[clust$tree_col$order,]), border_color=NA, cluster_rows=F, 
         fontsize=17, fontsize_row = 21, main="WT") +
  theme(plot.title = element_text(hjust = 0.7)) 
#dev.off()
```

Correlation between subsystem scores and protein abundance
```{r}
subsystem.cor <- subsystem.correlation(group.info = c("highest cell count", "highest viability", "highest amount", "lowest amount", "lowest viability", "lowest cell count"), scoreMatrix = delta_subsystemScore) 
```


** Per reviewers request: redo heatmap, but order rows by yield
```{r}
df.tmp <- t(delta_subsystemScore[,-96])
df.yield.tmp <- dplyr::select(CHO96, c("genes", "Total.purified.protein_stock1.stock2_microg"))[-96 ,] %>%
  arrange(desc(Total.purified.protein_stock1.stock2_microg))
df.reorder <- df.tmp[match(df.yield.tmp$genes, rownames(df.tmp)) ,]

## Add row annotations according to correlation

## add node colors
colfunc <- colorRampPalette(c("red","yellow","cyan","blue"))
#colfunc.node <- colorRampPalette(c("#9E3D22", "gray88", "#2B5C8A"))
systemCor.col <- data.frame(cor=round(seq(0, 0.5, 0.01), digits=2), color=colfunc(length(seq(0, 0.5, 0.01))))
idx <- match(round(subsystem.cor$rho, digits = 2), systemCor.col$cor)
identical(round(subsystem.cor$rho, digits = 2), systemCor.col$cor[idx])
annotation.col <- data.frame(correlation=systemCor.col$color[idx]) %>% `rownames<-`(subsystem.cor$gene)
ann.col$correlation= systemCor.col$color[idx] %>% setNames(systemCor.col$color[idx])



## plot pheatmap  
png("Figures/FigureS4.png", width=550, height=800)
clust <- pheatmap(df.reorder, fontsize=18, fontsize_col=15,
                               cluster_rows = F, cluster_cols = T, annotation_row = row.ann.95, annotation_col = annotation.col, gaps_col=T, annotation_colors = ann.col, show_rownames=F, color=myColor, breaks=myBreaks, annotation_legend = T)
dev.off()

rm(df.tmp)
rm(df.yield.tmp)
rm(df.reorder)


color.bar(colfunc(length(seq(0, 0.5, 0.01))), min=0, max=0.5, title="", ntick=3)

```


############################################################
Correlations within each cluster
############################################################
Given a secretory pathway phenotype (belonging to a cluster... respond/stressed out in a similar fashion, are there certain genes that can rescue protein secretion?)


Calculate correlations within each cluster
```{r}
cluster.gene.corr <- lapply (c(1:3), geneCorr.cluster) %>% setNames(c(1:3))

### add column for sybsystem
clust3.cor.suppMat <- cluster.gene.corr$`3`[order(-rho),] %>% 
  left_join(secM_TPM.map[, c("choSymbol", "Subsystem")], by=c("gene"="choSymbol")) %>% dplyr::rename(FDR=padj) %>%
  distinct()
openxlsx::write.xlsx(clust3.cor.suppMat, "Output/Cluster3_cor.xlsx")

## summary of FDR and Rho
lapply(cluster.gene.corr, function(x){summary(x$padj)})
lapply(cluster.gene.corr, function(x){summary(x$rho)})
```
Note that cluster 3 is the only cluster with significant correlations

```{r}
sigClust.3 <- sig.clust.cor(clusterID='3', FDR.thresh=0.1, rho.thresh=0.6)
#sigClust.2 <- sig.clust.cor(clusterID='2', FDR.thresh=0.1, rho.thresh=0.6)
#sigClust.1 <- sig.clust.cor(clusterID='1', FDR.thresh=0.1, rho.thresh=0.6)

saveRDS(sigClust.3$sig.plot, "Figures/R_plots/Figure4c.RDS")
```

Plot distribution of protein titer for each cluster
```{r}
df <- clust.samples[clust.samples$cluster != 4 ,] %>% mutate(cluster2=cluster)
df$cluster2[df$cluster==1] <- 2
df$cluster2[df$cluster==2] <- 1

mu <- plyr::ddply(df, "cluster", summarise, grp.mean=mean(yield))
ggplot(df, aes(x=yield, fill=as.factor(cluster))) + geom_density(alpha=0.4) +
  geom_vline(data=mu, aes(xintercept=grp.mean, color=as.factor(cluster)),linetype="dashed", show.legend = F) + 
  labs(fill = "Cluster") + xlab("abundance (μg)") + theme_bw() + theme(text=element_text(size=15))

#Note: 2 is 1, 1 is 2
data.clust <- df %>% left_join(CHO96[, c("sample_ID", "amount_microgram")]) %>% data.table()
```

############################################################
Quantify var explained in Cluster 3
############################################################
```{r}
select.genes <- c("Alg12","Derl2", "Rpn1", "Rpn2", "Ddost")
select.samples <- clust.samples$sample_ID[clust.samples$cluster %in% 3]
TPM.subset <- subset(TPM.log.secM, select=select.samples) %>% t() %>% as.data.frame()
response.var <- CHO96[CHO96$sample_ID %in% select.samples, c("sample_ID", "Total.purified.protein_stock1.stock2_microg")]
identical(rownames(TPM.subset), response.var$sample_ID)
TPM.subset <- subset(TPM.subset, select=select.genes) %>%
  mutate(response.var=log10(response.var$Total.purified.protein_stock1.stock2_microg + 0.001))

corrplot::corrplot(cor(TPM.subset), method="circle", addCoef.col = 1, tl.cex = 1.2)

fit.subset <- dplyr::select(TPM.subset, c("Alg12", "Derl2", "response.var"))
fit <- stats::lm(response.var ~ ., data=fit.subset )

summary(fit)
car::avPlots(fit, marginal.scale=F, id=F)
```


############################################################
Are clusters characterized by specific protein features
############################################################
load protein fetures used in ML
```{r}
protFeat <- readRDS("Output/protFeatures96.RDS") %>% left_join(clust.samples) #%>% mutate(cluster=as.factor(cluster))

protFeat.all <- readRDS("../ProteinFeatures/Output/Tegel2020_allFeatures.RDS")
protFeat.subset <- protFeat.all$features_allSamples %>% 
  right_join(clust.samples, by=c("human_symbol"="gene")) %>%
  dplyr::filter(cluster != 4) %>%
  mutate(cluster=as.factor(cluster))
```

Anova
```{r}
anova.feat <- dplyr::select(protFeat.subset, "abundance_GTEx_Adipose":cluster)
colnames(anova.feat) <-str_replace(names(anova.feat), "%", "percent")
colnames(anova.feat) <- str_replace(names(anova.feat), "-", "_")

# run anova for ever protein feature
res.anova <- sapply(names(anova.feat)[-219], anova.fxn) %>%
  lapply(function(x) {x$`Pr(>F)`[1]})

summary(unlist(res.anova))
View(sort(unlist(res.anova)))
```
Note: no defining features in each cluster

############################################################
Compare secM exp low vs high prod in each cluster
############################################################

two-sided welch test comparing low vs high producers in each cluster
```{r}
welch.cluster(1, TPM.log.secM, alt="two.sided")[padj <=0.1 ,] %>%
  mutate(LFC=mean.x-mean.y)
welch.cluster(2, TPM.log.secM, alt="two.sided")[padj <=0.1 ,]
welch.cluster(3, TPM.log.secM, alt="two.sided")[padj <=0.1 ,]
```

Look at GPI PTM in Cluster 1
```{r}
# which rProteins in Cluster 1 contain GPI anchor PTM -- NONE
dplyr::select(protFeat.subset, c("human_symbol", "sequence_PSIM_GPI", "sequence_PTM.detailed_Lipidation__GPI.anchor")) %>%
  left_join(data.clust, by=c("human_symbol"="gene")) %>% na.omit() #%>%View()

# load list of host HCP proteins in CHO [10.1038/srep28547]
GPI.HCP <- openxlsx::read.xlsx("Data/protFeatures/IoscaniJimenez_2016.xlsx", sheet="ST10 - GAG and GPI", startRow=3, colNames=T, check.names=T) %>% 
  dplyr::select(c("Entry.Name", "GPI.Anchor")) %>%
  dplyr::filter(GPI.Anchor==1) %>% 
  mutate(geneName=gsub("_.*","",Entry.Name)) %>%
  left_join(dplyr::select(list.TPM_pheno_map$CHO.human_map, c("choSymbol", "humanSymbol")), by=c("geneName"="humanSymbol"))

# load other list of GPI anchor proteins from secRecon (need to map to gene names)
CHOmap <- fread("ConversionTable/geneInfo_cho_simple.tsv") %>% # map entrezID to gene name
  mutate(GeneID=as.character(GeneID)) %>%
  # map uniprotID to entrezID
  inner_join(openxlsx::read.xlsx("ConversionTable/CHOProtein.xlsx"), by=c("GeneID"="ENTREZ_GENE"))
PSIM.GPI <- openxlsx::read.xlsx("Data/protFeatures/PSIM_matrix.xlsx", sheet="Chinese hamster") %>%
  filter(GPI==1) %>% 
  left_join(dplyr::select(CHOmap, c("UNIPROTKB", "Symbol")), by=c("Entry"="UNIPROTKB"))
  
# combine
GPI.symbol <- c(GPI.HCP$choSymbol, PSIM.GPI$Symbol) %>% na.exclude()

# extract expression (TPM) of the CHO GPI 
TPM.GPI <- dplyr::select(list.TPM_pheno_map$TPM.log, -sds) %>%
  subset(gene %in% GPI.symbol) %>% na.omit() %>% distinct()

#two-sided welch test comparing the expression of GPI-anchor poteins in low vs high producers
exp.data= tibble::column_to_rownames(TPM.GPI, "gene") %>% as.matrix()

welch.cluster(1, exp.data, alt="two.sided")[padj <=0.1 ,] #cluster 1
welch.cluster(2, exp.data, alt="two.sided")[padj <=0.1 ,] #cluster 2
welch.cluster(3, exp.data, alt="two.sided")[padj <=0.1 ,] #cluster 3


```
Note: No differences in GPI-anchor protiens. The difference in GPI-anchor synthesis gene is prob artifact of being very lowly expressed.




